<style type="text/css">

/* Table of content - navigation */
div#TOC li {
    list-style:none;
    background-color:skyblue;
    background-image:none;
    background-repeat:none;
    background-position:0;
    font-family: Arial, Helvetica, sans-serif;
    color: darkred;
}

h1.title {
  font-size: 24px;
  color: DarkRed;
  text-align: center;
}

h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h1 { /* Header 3 - and the author and data headers use this too  */
    font-size: 22px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: center;
}
h2 { /* Header 3 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h3 { /* Header 3 - and the author and data headers use this too  */
    font-size: 15px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h4 { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: left;
}
</style>


```{r setup, include=FALSE}
# code chunk specifies whether the R code, warnings, and output 
# will be included in the output files.
if (!require("tidyverse")) {
   install.packages("tidyverse")
   library(tidyverse)
}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("cowplot")) {
   install.packages("cowplot")
   library(cowplot)
}
if (!require("latex2exp")) {
   install.packages("latex2exp")
   library(latex2exp)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("gapminder")) {
   install.packages("gapminder")
   library(gapminder)
}
if (!require("png")) {
    install.packages("png")             # Install png package
    library("png")
}
if (!require("RCurl")) {
    install.packages("RCurl")             # Install RCurl package
    library("RCurl")
}
if (!require("colourpicker")) {
    install.packages("colourpicker")              
    library("colourpicker")
}
if (!require("gganimate")) {
    install.packages("gganimate")              
    library("gganimate")
}
if (!require("gifski")) {
    install.packages("gifski")              
    library("gifski")
}
if (!require("magick")) {
    install.packages("magick")              
    library("magick")
}
if (!require("grDevices")) {
    install.packages("grDevices")              
    library("grDevices")
}
if (!require("jpeg")) {
    install.packages("jpeg")              
    library("jpeg")
}
if (!require("VGAM")) {
    install.packages("VGAM")              
    library("VGAM")
}
if (!require("MASS")) {
    install.packages("MASS")              
    library("MASS")
}
if (!require("nnet")) {
    install.packages("nnet")              
    library("nnet")
}
if (!require("cluster")) {
    install.packages("cluster")              
    library("cluster")
}
# knitr::opts_knit$set(root.dir = "C:/Users/75CPENG/OneDrive - West Chester University of PA/Documents")
# knitr::opts_knit$set(root.dir = "C:\\STA490\\w05")

knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE,
                      comment = NA)

x <- getURL("https://jessgorr01.github.io/jGorrCV/wk3/w03-penguins.csv")
data <- read.csv(text = x)
```


# Jessica Gorr

Data Management Assignment: Penguin Dataset

## Data Management

This dataset "penguin.csv" contains multiple variables about penguins such as: species, location, and various measurement. Not all the data is complete so before analyzing any data, it is important to clean up and arrange the data in a useful way.

The following code chunk deletes any cases that have missing values.
```{r}
dataCleaned <- data[complete.cases(data),]
```

The next code chunk will sort out and include only the penguins with the following conditions:
1. Include Adelie penguins and Gentoo penguins from the Biscoe and Torgersen islands.
2. Include only penguins with body_mass_g less than 5000 grams but more than 3500 grams.
3. Rescale body_mass_g by dividing 4000 and rename it as BMI = body_mass_g / 4000.
4. Exclude variables X(observation ID), sex, year, and body_mass_g from the above subset.

```{r}

dataFiltered <- dataCleaned %>%
              filter((species == "Adelie" | species =="Gentoo") & (island == "Biscoe" | island == "Torgersen") )%>%
  mutate(BMI = body_mass_g/4000)%>% 
             filter(body_mass_g < 5000 & body_mass_g > 3500)%>%
  dplyr::select(-X,-sex,-year,-body_mass_g)
```
## Scatterplot

Now a simple scatter-plot made with the `plot()` function featuring the two variables Bill_length_mm and Flipper_length_mm from the penguins data set. Since the data is sub setted from the original, there are now less observations

```{r}
library(jpeg)
iris.img <- "https://jessgorr01.github.io/jGorrCV/wk3/penguin.jpeg"
my.iris <- readJPEG(getURLContent(iris.img))
raster.iris <- as.raster(my.iris) 


# Use the code in the precious section
flipper.length = dataFiltered$flipper_length_mm
Bill.length = dataFiltered$bill_length_mm
species = dataFiltered$species
size.body = dataFiltered$BMI
## identifying the ID of Virginica
Adelie.id = which(species=="Adelie")  # value are case sensitive!
Gentoo.id = which(species=="Gentoo")
## color code
 #col.code = c(alpha("red",0.5),alpha("blue",0.5),alpha("cyan",0.5))
## making an empty plot: type = "n" ==> no point
plot(Bill.length, flipper.length, main = "Relationship of Flipper and Bill Length Amounst all Species of Penguins", type = "n",  xlab="Bill Length (mm)", ylab="Flipper Length (mm)")
## change the point size based on their average of sepal length and width
points(Bill.length[Adelie.id], flipper.length[Adelie.id], 
       pch = 19, col = "purple", cex = size.body[Adelie.id])
points(Bill.length[Gentoo.id], flipper.length[Gentoo.id], 
       pch = 18, col = "navy", cex = size.body[Gentoo.id])
legend("topleft", c("Adelie", "Gentoo"), 
                  col=c("purple", "navy"),
                  pch=c(19, 18))

abline(lm(flipper.length[Adelie.id]~ Bill.length[Adelie.id]), col = "purple")
abline(lm(flipper.length[Gentoo.id]~ Bill.length[Gentoo.id]), col = "navy")


## various annotations
#specify the position of the image through bottom-left and top-right coords
rasterImage(my.iris, 
            xleft=45, xright=50, 
            ybottom=180, ytop=200)

```